Does the COMPAS Needle Always Point Towards Equity? Finding Fairness in the COMPAS Risk Assessment Algorithm: A Case Study

### 2022-05-04 Paper_Draft2/Draft2.Rmd

With our baseline model, we ran the four different group fairness metrics we chose and compared the results (pictured in Figre \ref{fig:baseline metrics}).

The statistical parity difference is -0.14. This indicates that there is a large difference between white and Black defendants regarding whether or not they recidivate. The algorithm unfairly benefits white defendants over Black defendants.

@@

Through our methods, we concluded that the pre-processing and post-processing methods mitigate bias and increase fairness most successfully. The in-processing technique was effective at mitigating bias, but did not result in a mathematically fair model. However, these techniques only treat the symptoms of racial bias in the justice system, as opposed to addressing the root cause, systemic racism. We need to question why these flawed algorithms exist and why they can be used to determine someone’s fate in the justice system. While we may have methods for reducing bias algorithms such as COMPAS, the focus should be on addressing the over-policing and over-criminalization of Black communities that result in biased data and algorithms.

@@

```{r word count, include=F}
wordcountaddin:::text_stats()
```

>>>>

With our baseline model, we ran the four different group fairness metrics we chose and compared the results (pictured in Figure \ref{fig:baseline metrics}).

Statistical parity difference is -0.14. This indicates that there is a large difference between white and Black defendants regarding whether or not they recidivate. The algorithm unfairly benefits white defendants over Black defendants.

@@

Through our methods, we concluded that the pre-processing and post-processing methods mitigate bias and increase fairness most successfully. The in-processing technique was effective at mitigating bias, but did not result in a mathematically fair model. However, these techniques only treat the symptoms of racial bias in the justice system, as opposed to addressing the root cause, systemic racism. We need to question why these flawed algorithms exist and why they can be used to determine someone’s fate in the justice system. While we may have methods for reducing biased algorithms such as COMPAS, the focus should be on addressing the over-policing and over-criminalization of Black communities that result in biased data and algorithms.

@@

```{r word count, warning=FALSE, message=FALSE, echo=FALSE}
wordcountaddin:::text_stats()
```

### 2022-05-04 Paper_Draft2/Draft2.tex

Through our methods, we concluded that the pre-processing and
post-processing methods mitigate bias and increase fairness most
successfully. The in-processing technique was effective at mitigating
bias, but did not result in a mathematically fair model. However, these
techniques only treat the symptoms of racial bias in the justice system,
as opposed to addressing the root cause, systemic racism. We need to
question why these flawed algorithms exist and why they can be used to
determine someone's fate in the justice system. While we may have
methods for reducing bias algorithms such as COMPAS, the focus should be
on addressing the over-policing and over-criminalization of Black
communities that result in biased data and algorithms.

@@

>>>>

Through our methods, we concluded that the pre-processing and
post-processing methods mitigate bias and increase fairness most
successfully. The in-processing technique was effective at mitigating
bias, but did not result in a mathematically fair model. However, these
techniques only treat the symptoms of racial bias in the justice system,
as opposed to addressing the root cause, systemic racism. We need to
question why these flawed algorithms exist and why they can be used to
determine someone's fate in the justice system. While we may have
methods for reducing biased algorithms such as COMPAS, the focus should
be on addressing the over-policing and over-criminalization of Black
communities that result in biased data and algorithms.

@@

\begin{longtable}[]{@{}lll@{}}
\toprule
Method & koRpus & stringi\tabularnewline
\midrule
\endhead
Word count & 4719 & 4537\tabularnewline
Character count & 29868 & 29825\tabularnewline
Sentence count & 280 & Not available\tabularnewline
Reading time & 23.6 minutes & 22.7 minutes\tabularnewline
\bottomrule
\end{longtable}
